---
title: "Chord Diagrams"
output: html_notebook
---

# load packages

```{r, message=FALSE}
library(circlize)
```

# create data

```{r}
data <- data.frame(
  factor = sample(letters[1:8], 1000, replace = TRUE),
  x = rnorm(1000),
  y = runif(1000)
)

data
```

# First circos

Three steps must be followed:

```{r}
# 1 Initialise
circos.initialize(
  factors = data$factor,
  x = data$x
)

# 2 Build the regions
circos.trackPlotRegion(
  factors = data$factor,
  y = data$y,
  panel.fun = function(x, y) {
    circos.axis()
  }
)


# 3 Add points
circos.trackPoints(
  data$factor,
  data$x,
  data$y,
  # point type:
  pch = 16, #simple point
  # character expansion:
  cex = 0.5,
  col = 'cornflowerblue'
)

```
# Chord Diagram and Adjacency Matrix

Proceed to create a list of connections between 20 origin nodes and 5 destinations:

```{r}
# define a random interval of points
set.seed(42)
nums <- sample(
  # interval:
  c(1:200),
  # number drawn:
  100,
  # replacement
  replace = TRUE
  )

# turn the numbers into a matrix
df <- matrix(
  nums,
  ncol = 5
)

# change rownames and colnames of the data
rownames(df) <- paste0('origin_', 1:20)     # can also use seq(1,20)
colnames(df) <- paste0('destination_', 1:5) # can also use seq(1,5)

head(df)
```

And making a chord diagram is very fast:

```{r}
chordDiagram(df)
```

Note: the data must be an adjacency matrix.

# Advanced Chord Diags

```{r, message=FALSE}
library(tidyverse)
library(viridis)
```


Let's get back to the migration data:

```{r}
# store the url data
url <- 'https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/13_AdjacencyDirectedWeighted.csv'

# read the data
migrations <- read.table(url)

# have the same rownames and colnames
colnames(migrations) <- rownames(migrations)

migrations
```

Still using `circlize` in three steps, but we can use it with `long` data as well:

```{r}
data_long <-
  migrations %>%
  rownames_to_column() %>%
  rename(origin = rowname) %>%
  pivot_longer(
    # data to use
    !origin,
    # gather col names to:
    names_to = 'destination',
    # gather col rows to:
    values_to = 'n'
  ) %>%
  filter(n > 0)

data_long
```

Then we initialise the steps:

```{r}
# not compatible in tidy grammar, sigh

circos.clear() # reset params

circos.par( # set a bunch of stuff
  start.degree = 90,
  gap.degree = 4,
  track.margin = c(-0.1, 0.1),
  points.overflow.warning = FALSE
)

# set plot parameters:
par(mar = # margins
      rep(0,4)) # 0 repeated 4 times

# color palette using 'viridis' library:
mycolor <- viridis(10, alpha = 1, begin = 0, end = 1, option = "D")
mycolor <- mycolor[sample(1:10)]
```

Then create the base plot:

```{r}
# base plot
chordDiagram(
  x = data_long, 
  grid.col = mycolor,
  transparency = 0.25,
  directional = 1,
  direction.type = c("arrows", "diffHeight"), 
  diffHeight  = -0.04,
  annotationTrack = "grid", 
  annotationTrackHeight = c(0.05, 0.1),
  link.arr.type = "big.arrow", 
  link.sort = TRUE, 
  link.largest.ontop = TRUE)

# text and axis
circos.trackPlotRegion(
  track.index = 1, 
  bg.border = NA, 
  panel.fun = function(x, y) {
    
    xlim = get.cell.meta.data("xlim")
    sector.index = get.cell.meta.data("sector.index")
    
    # Add names to the sector. 
    circos.text(
      x = mean(xlim), 
      y = 3.2, 
      labels = sector.index, 
      facing = "bending", 
      cex = 0.8
      )

    # Add graduation on axis
    # circos.axis(
    #   h = "top", 
    #   major.at = seq(from = 0, to = xlim[2],
    #                  by = ifelse(test = xlim[2]>10, yes = 2, no = 1)), 
    #   minor.ticks = 1, 
    #   major.tick.length = 0.5,
    #   labels.niceFacing = FALSE)
  }
)

```
