---
title: "Sankey Diagram"
output: html_notebook
---

# Packages required

```{r, message=FALSE}
# for ggplot2 and data manipulation
library(tidyverse)

# for the diagram
library(networkD3)
```

# Load Data

```{r}
weburl = 'https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/13_AdjacencyDirectedWeighted.csv'

data <-
  read.table(weburl, header = TRUE)

data
```

## Some wrangling

```{r}
data_long <-
  data %>%
  # transform indices into a column on its own
  rownames_to_column() %>%
  pivot_longer(
    -rowname, # this is the name of the col created via 'rownames_to_column()'
    names_to = 'key',
    values_to = 'value'
  ) %>%
  filter(value > 0) %>%
  mutate(
    # remove the dot between the values
    key = str_replace_all(key, '[\\.]', ' '),
    # a little trick: add one empty space at the end of each value
    key = paste0(key, ' ')
  ) %>%
  rename(
    origin = rowname,
    destination = key,
  )

data_long
```

# Set up

From the flows data we need to create the nodes. We use `unique` to return a dataframe with only the destinations, with no repetition.

```{r}
nodes <-
  data.frame(
    name = c(data_long$origin, data_long$destination)
  ) %>% unique()

nodes
```

Data appears twice because we added a blank space at the end of `destination`: in this way, we will have the data displayed in two columns even when the starting point is the same as the end point.

Then we match each name in the `data_long` with the corresponding one in `nodes`:

```{r}
data_long %>%
  mutate(
    # we need to use `match`, not `as.numeric`!
    id_origin = match(data_long$origin, nodes$name) - 1,
    id_destination = match(data_long$destination, nodes$name) -1
  ) %>%
  as.data.frame %>%
  sankeyNetwork(
  Nodes = nodes,
  Source = 'id_origin',
  Target = 'id_destination',
  Value = 'value',
  NodeID = 'name',
  # sinksRight = FALSE,
  nodeWidth=40, fontSize=13, nodePadding=20
)

```


